### 2013-09-25 evaluation/paper.tex
bring experiments section into some sort of shape... at least for exact inference.

\subsection{Experiments using Exact Inference}
In this section, we present results on multi-class classification, sequence
labeling and multi-label prediction. In these
tasks, exact inference is possible, and we can compute the exact objective of
the various algorithms easily. We use the dynamic programming algorithm
implemented in \textsc{OpenGM} for experiments for the tree and chain models.
For the full pairwise multi-label model, we use branch-and-bound together with the %FIXME AD3 solver description?
We found that using the linear programming relaxation on the scene dataset is
often tight, and there was no need to resort to approximate inference. Inference
on the yeast dataset was more complex, and on the verge of being non-practical.
We use this dataset as an example for very cosly inference.

We start our experiments with the MNIST dataset.
The convergence of the primal objective and training set loss in terms of
passes over the training set is shown in \Figref{mnist_iterations}.
Several trends can be observed: the $n$-slack cutting plane algorithms
converge fastest, with the version that recomputes the QP at every step leading
the race. The algorithms are closely followed by the caching $1$-slack cutting
plane algorithm.  They are followed by the significantly slower SZLJSP
algorithm, and finally the non-caching $1$-slack cutting plane algorithm and
subgradient descent.  These results are intuitive, as they reflect ``how much
work'' each algorithm does for each loss-augmented prediction step. More work
towards the objective leads to faster convergence.  This ``more work'' is
quantified in \Figref{mnist_times}.  Clearly the $1$-slack algorithms to ``to
much'' work, leading to very slow convergence.  Also, caching does not speed up
learning learning on this dataset, as loss-augmented prediction is as fast as
querying the cache.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/letters_small_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/letters_small_log_time}
    \end{tabu}
\caption{%
   Convergence of the primal objective on OCR-small. 
\figlabel{ocr_small_results}}
\end{figure}

We now consider datasets with non-trivial inference.
The results for OCR-small are shown in \Figref{ocr_small_results}.
The trends seen in the figures are similar to what we saw for MNIST above.
Again, even though inference is non-trivial, the non-caching $1$-slack
algorithm is slower than the caching one. Again, the plots against passes over the training set
do not reflect the runtime complexity of the $n$-slack approach.

\begin{figure}
    \begin{center}
    \includegraphics[width=.5\linewidth]{evaluation/images/letters_small_dual}
\end{center}
\caption{%
   Convergence of the primal objective on OCR-small. 
\figlabel{ocr_small_dual}}
\end{figure}

There is one interesting phenomenon to be seen here: The cutting-plane algorithms stop immediately
when reaching the desired suboptimality of $0.1$. The SZLJSP algorithm, on the other hand, keeps
on learning much longer---even longer than it took to get to the desired suboptimality. This is caused
by a much looser bound given by the dual. Remember that the stopping criterion for both cutting-plane 
and SZLJSP are given by the duality gap. It seems that while the primal objective converges much
faster in the SZLJSP algorithm than in the non-caching $1$-slack algorithm, the dual does not.
This means that in for this practical experiment, the non-caching $1$-slack algorithm was \emph{faster}
in guaranteeing the desired suboptimality.
\Figref{ocr_small_dual} illustrates this point by ploting primal and dual
objectives for the $1$-slack cutting plane and the SZLJSP algorithms.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/letters_big_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/letters_big_log_time}
    \end{tabu}
\caption{%
   Convergence of the primal objective on OCR-large. 
\figlabel{ocr_large_results}}
\end{figure}

The results for the OCR-large dataset are shown in \Figref{ocr_large_results}. It was not feasible to run the $n$-slack
algorithm when solving the QP at every step here.
The trends are the same as for the smaller dataset: $n$-slack and caching $1$-slack cutting plane are very fast
in terms of passes through the dataset, but $n$-slack cutting plane is impractical slow
with respect to runtime. The SZLJSP algorithm converges fast intially, but the $1$-slack cutting plane
is faster at high precision and faster in certifying the desired duality gap.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/scene_full_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/scene_full_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}
    \end{tabu}
\caption{%
   Convergence of the primal objective on the multi-label datasets. From top to
   bottom: the scene dataset using a tree model, the scene dataset using a full
   model, the yeast dataset using a tree model, and the yeast dataset using a
   full model.
\figlabel{multi-label-results}}
\end{figure}

Next, we consider the multi-label task.  The results for the yeast and scene
datasets are shown in \Figref{multi-label-results}, respectively. The caching
$1$-slack algorithm is much faster than the non-caching one on both datasets.
While the two variants of the $n$-slack algorithm are equally fast with respect
to passes over the training data, solving the QP every 100 steps is among the
fastest algorithms, while solving the QP every step is among the slowest.

The caching $1$-slack algorithm is the fastest algorithm to achieve the desired
primal suboptimality.  On the scene dataset, the other cutting-plane base
algorithms are second fastest, while they lag behind the stochastic algorithms
on the larger yeast dataset. Again, initial convergence is faster for the
stochastic algorithms.

@@

\section{Summary}
From the experiments above, it is clear there is currently \emph{no single best algorithm}.
However, there are some clear trends:
\begin{itemize}
    \item SZJLSP always converges faster than SSGD, with each iteration having
        the same time complexity.  This indicates that SZJLSP should be
        preferred over SSGD, apart from settings with strong memory
        constraints.
    \item In nearly all experiments, caching significantly improved performance
        of the 1-slack cutting plane algorithm.
    \item Initially SZJLSP and even SSGD converge faster than the 1-slack
        cutting plane algorithm, while the 1-slack cutting plane algorithm was
        faster in achieving high precision solutions and guaranteeing low
        duality gap.
    \item The n-slack cutting plane algorithm was often fastest to converge in
        terms of passes over the training set, but often slow in terms of
        runtime. The caching $1$-slack algorithm converged faster in terms of
        cpu-time on almost all datasets.
\end{itemize}
There are however some caveats with our analysis, in particular with respect to
the cutting plane solvers. The performance of the cutting plane algorithms
depends not only on the performance of the QP solver used (in our case
quadratic cone programming), but also on heuristics for pruning variables from
the QP, caching constraints (for the $1$-slack algorithm) and deciding how
often to solve the QP (for the $n$-slack algorithm). Better heuristics could
have a beneficial effect on these algorithms. On the other hand, it is
remarkable how competative the SZLJSP algorithm is, without the need for any
such implementation tricks. This makes the SZJLSP algorithm very attractive for
the practitioner who does not want to spend much time tuning the
implementation.

We do not consider SSGD using averaging here, but found it so produce similar
results in tentative experiments. In particular, it was never faster than SZLJSP.

With respect to approximate learning, we found that the inference algorithm has
a stronger impact on the performance than the learning algorithm.
While the SSGD and SZJLSP algorithms have somewhat lower training error on the snakes than
the cutting-plane algorithm when using fusion-move inference, they, too, fail at solving the task.
Using the linear programming relaxation, on the other hand, all algorithms are able to solve the
task near perfect.
For the segmentation datasets, the approximate inference seem to be less of a
problem, as most algorithms achieve the desired approximate duality gap (which
is not the actual duality gap, see Chapter~\ref{ch:exact_learning} for a
discussion).
The large label space made using the linear programming relaxation intractable,
though, and we can not draw a final conclusion here. In the next chapter,
Chapter~\ref{ch:exact_learning}, we will demonstrate how we can learn exactly,
even in this setting.

>>>>

\subsection{Experiments using Exact Inference}
In this section, we present results on multi-class classification, sequence
labeling and multi-label prediction. In these tasks, exact inference is
possible, and we can compute the exact objective of the various algorithms
easily. We use the dynamic programming algorithm implemented in \textsc{OpenGM}
for experiments for the tree and chain models.
For the full pairwise multi-label model, we use branch-and-bound together with the %FIXME AD3 solver description?
We found that using the linear programming relaxation on the scene dataset is
often tight, and there was no need to resort to approximate inference. Inference
on the yeast dataset was more complex, and on the verge of being non-practical.
We use this dataset as an example for very costly inference.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/mnist}&%
    \includegraphics[width=\linewidth]{evaluation/images/mnist_loss}
    \end{tabu}
\caption{%
   Convergence of the primal suboptimality and training set loss on MNIST. 
\figlabel{mnist_curves}}
\end{figure}

We start our experiments with the MNIST dataset.
The convergence of the primal suboptimality and training set loss in terms of
passes over the training set is shown in \Figref{mnist_curves}. While inference
is trivial, this is by far the largest dataset, making the $n$-slack algorithms
not feasible. We see that the stochastic algorithms are much faster than the
cutting plane algorithm, in particular initially, even in terms of iterations,
with the SZLJSP is clearly leading the way. Interestingly, the $1$-slack algorithm
catches up near the desired suboptimality. Looking at the training set loss,
it is clear that the high precision we demanded was not necessary, and we could have
terminated the stochastic algorithms much earlier.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/letters_small_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/letters_small_log_time}
    \end{tabu}
\caption{%
   Convergence of the primal suboptimality on OCR-small. 
\figlabel{ocr_small_results}}
\end{figure}

We now consider datasets with non-trivial inference.
The results for OCR-small are shown in \Figref{ocr_small_results}.
Considering the plot against passes over the dataset on the left, several
trends can be observed: the $n$-slack cutting plane algorithms converge
fastest, with the version that recomputes the QP at every step leading the
race. The algorithms are closely followed by the caching $1$-slack cutting
plane algorithm.  They are followed by the significantly slower SZLJSP
algorithm, and finally the non-caching $1$-slack cutting plane algorithm and
subgradient descent.  These results are intuitive, as they reflect ``how much
work'' each algorithm does for each loss-augmented prediction step. More work
towards the objective leads to faster convergence.  This ``more work'' is
quantified on the right hand side, where the suboptimality is plotted against time.
Clearly the $1$-slack algorithms to ``to much'' work, leading to very slow
convergence.  Also, caching does not speed up learning learning on this
dataset.

%The trends seen in the figures are similar to what we saw for MNIST above.
%Again, even though inference is non-trivial, the non-caching $1$-slack
%algorithm is slower than the caching one. Again, the plots against passes over the training set
%do not reflect the runtime complexity of the $n$-slack approach.

\begin{figure}
    \begin{center}
    \includegraphics[width=.5\linewidth]{evaluation/images/letters_small_dual}
\end{center}
\caption{%
    Primal and dual objectives for SZLJSP and $1$-slack cutting plane on OCR-small.
\figlabel{ocr_small_dual}}
\end{figure}

We want to highlight an interesting phenomenon here: The cutting-plane algorithms stop immediately
when reaching the desired suboptimality of $0.1$. The SZLJSP algorithm, on the other hand, keeps
on learning much longer---even longer than it took to get to the desired suboptimality. This is caused
by a much looser bound given by the dual. Remember that the stopping criterion for both cutting-plane 
and SZLJSP are given by the duality gap. It seems that while the primal objective converges much
faster in the SZLJSP algorithm than in the non-caching $1$-slack algorithm, the dual does not.
This means that in this practical experiment, the non-caching $1$-slack algorithm was \emph{faster}
in guaranteeing the desired suboptimality.
\Figref{ocr_small_dual} illustrates this point by plotting primal and dual
objectives for the $1$-slack cutting plane and the SZLJSP algorithms.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/letters_big_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/letters_big_log_time}
    \end{tabu}
\caption{%
   Convergence of the primal suboptimality on OCR-large. 
\figlabel{ocr_large_results}}
\end{figure}

The results for the OCR-large dataset are shown in \Figref{ocr_large_results}. It was not feasible to run the $n$-slack
algorithm when solving the QP at every step here.
The trends are the same as for the smaller dataset: $n$-slack and caching $1$-slack cutting plane are very fast
in terms of passes through the dataset, but $n$-slack cutting plane is impractical slow
with respect to runtime. The SZLJSP algorithm converges fast initially, but the $1$-slack cutting plane
is faster at high precision and faster in certifying the desired duality gap.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/scene_full_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/scene_full_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/yeast_full}&%
    \includegraphics[width=\linewidth]{evaluation/images/yeast_full_time}
    \end{tabu}
\caption{%
   Convergence of the primal objective on the multi-label datasets. From top to
   bottom: the scene dataset using a tree model, the scene dataset using a full
   model, the yeast dataset using a tree model, and the yeast dataset using a
   full model.
\figlabel{multi-label-results}}
\end{figure}

Next, we consider the multi-label task.  The results for the yeast and scene
datasets are shown in \Figref{multi-label-results}. The caching $1$-slack
algorithm is much faster than the non-caching one on both datasets and both
graph structures.
While the two variants of the $n$-slack algorithm are equally fast with respect
to passes over the training data, solving the QP every 100 steps is among the
fastest algorithms, while solving the QP every step is among the slowest.
We applied this method only to the smaller scene dataset, as applying it on the
yeast dataset was not practicable.

The caching $1$-slack algorithm is the fastest algorithm to achieve the desired
primal suboptimality for tree-structured graphs, while it is out-performed by
the $n$-slack cutting plane algorithm for the full graphs. This is intuitive,
as the additional work the $n$-slack algorithm does becomes more valuable, the
longer the inference takes.

@@

\section{Summary}
From the experiments above, it is clear there is currently \emph{no single best algorithm}.
However, there are some clear trends:
\begin{itemize}
    \item SZLJSP always converges faster than SSGD, with each iteration having
        the same time complexity.  This indicates that SZJLSP should be
        preferred over SSGD, apart from settings with strong memory
        constraints.
    \item In nearly all experiments, caching significantly improved performance
        of the 1-slack cutting plane algorithm.
    \item Initially SZJLSP and even SSGD converge faster than the 1-slack
        cutting plane algorithm, while the 1-slack cutting plane algorithm was
        faster in achieving high precision solutions and guaranteeing low
        duality gap.
    \item The $n$-slack cutting plane algorithm was often fastest to converge in
        terms of passes over the training set, but often slow in terms of
        runtime. For problems with very slow inference, the $n$-slack algorithm
        might be a better choice than the others---if the interval between re-solving
        the QP is chosen right.
\end{itemize}
There are however some caveats with our analysis, in particular with respect to
the cutting plane solvers. The performance of the cutting plane algorithms
depends not only on the performance of the QP solver used (in our case
quadratic cone programming), but also on heuristics for pruning variables from
the QP, caching constraints (for the $1$-slack algorithm) and deciding how
often to solve the QP (for the $n$-slack algorithm). Better heuristics could
have a beneficial effect on these algorithms. On the other hand, it is
remarkable how competitive the SZLJSP algorithm is, without the need for any
such implementation tricks. This makes the SZLJSP algorithm very attractive for
the practitioner who does not want to spend much time tuning the
implementation.

We do not consider SSGD with averaging~\cite{lacoste2012simpler} here, but
found it so produce similar results in tentative experiments. In particular, it
never outperformed SZLJSP.

With respect to approximate learning, we found that the inference algorithm has
a stronger impact on the performance than the learning algorithm.
While the SSGD and SZLJSP algorithms have somewhat lower training error on the snakes than
the cutting-plane algorithm when using fusion-move inference, they, too, fail at solving the task.
Using the linear programming relaxation, on the other hand, all algorithms are able to solve the
task near perfect.
For the segmentation datasets, the approximate inference seem to be less of a
problem, as most algorithms achieve the desired approximate duality gap (which
is not the actual duality gap, see Chapter~\ref{ch:exact_learning} for a
discussion).
The large label space made using the linear programming relaxation intractable,
though, and we can not draw a final conclusion here. In the next chapter,
Chapter~\ref{ch:exact_learning}, we will demonstrate how we can learn exactly,
even in this setting.
